---
title: "Student Performance in Portugal"
author: "James Irwin, Cole Rowland, and Ryan Simonds"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
    html-table-processing: none
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---


```{r}
#| label: packages

library(tidyverse)
```

## Instructions

The [data](https://archive.ics.uci.edu/dataset/320/student+performance) for this
lab is provided by UC Irvine's Machine Learning Repository. This data involves
student achievement in secondary education at two Portuguese schools. The collected
data includes student demographic information, study habits, and extracurricular
activities. There are two datasets that outline performance in two different
subjects: math and Portuguese language.

::: {.callout-caution collapse="true"}
## Click here for detailed data description

Outline of each variable, as provided by UCI's Machine Learning Repository.

#### Attributes for both math and language datasets

1. **school** - student's school (binary)
   - GP: Gabriel Pereira
   - MS: Mousinho da Silveira

2. **sex** - student's sex (binary)
   - F: female
   - M: male

3. **age** - student's age (numeric)
   - From 15 to 22

4. **address** - student's home address type (binary)
   - U: urban
   - R: rural

5. **famsize** - family size (binary)
   - LE3: less or equal to 3
   - GT3: greater than 3

6. **Pstatus** - parent's cohabitation status (binary)
   - T: living together
   - A: apart

7. **Medu** - mother's education (numeric)
   - 0: none
   - 1: primary education (4th grade)
   - 2: 5th to 9th grade
   - 3: secondary education
   - 4: higher education

8. **Fedu** - father's education (numeric)
   - 0: none
   - 1: primary education (4th grade)
   - 2: 5th to 9th grade
   - 3: secondary education
   - 4: higher education

9. **Mjob** - mother's job (nominal)
   - teacher
   - health care related
   - services (e.g. administrative or police)
   - at_home
   - other

10. **Fjob** - father's job (nominal)
    - teacher
    - health care related
    - services (e.g. administrative or police)
    - at_home
    - other

11. **reason** - reason to choose this school (nominal)
    - home: close to home
    - reputation: school reputation
    - course: course preference
    - other

12. **guardian** - student's guardian (nominal)
    - mother
    - father
    - other

13. **traveltime** - home to school travel time (numeric)
    - 1: <15 min.
    - 2: 15 to 30 min.
    - 3: 30 min. to 1 hour
    - 4: >1 hour

14. **studytime** - weekly study time (numeric)
    - 1: <2 hours
    - 2: 2 to 5 hours
    - 3: 5 to 10 hours
    - 4: >10 hours

15. **failures** - number of past class failures (numeric)
    - n if 1 <= n < 3
    - otherwise 4

16. **schoolsup** - extra educational support (binary)
    - yes
    - no

17. **famsup** - family educational support (binary)
    - yes
    - no

18. **paid** - extra paid classes within the course subject (Math or Portuguese) (binary)
    - yes
    - no

19. **activities** - extra-curricular activities (binary)
    - yes
    - no

20. **nursery** - attended nursery school (binary)
    - yes
    - no

21. **higher** - wants to take higher education (binary)
    - yes
    - no

22. **internet** - Internet access at home (binary)
    - yes
    - no

23. **romantic** - with a romantic relationship (binary)
    - yes
    - no

24. **famrel** - quality of family relationships (numeric)
    - 1: very bad
    - 5: excellent

25. **freetime** - free time after school (numeric)
    - 1: very low
    - 5: very high

26. **goout** - going out with friends (numeric)
    - 1: very low
    - 5: very high

27. **Dalc** - workday alcohol consumption (numeric)
    - 1: very low
    - 5: very high

28. **Walc** - weekend alcohol consumption (numeric)
    - 1: very low
    - 5: very high

29. **health** - current health status (numeric)
    - 1: very bad
    - 5: very good

30. **absences** - number of school absences (numeric)
    - From 0 to 93

#### These grades are related with the course subject, Math or Portuguese

31. **G1** - first period grade (numeric)
    - From 0 to 20

32. **G2** - second period grade (numeric)
    - From 0 to 20

33. **G3** - final grade (numeric, output target)
    - From 0 to 20

#### Additional note

There are several (382) students that belong to both datasets. These student
can be identified by searching for identical attributes that characterize
each student. These features are:

- school
- sex
- age
- address
- famsize
- Pstatus
- Medu
- Fed
- Mjob
- Fjob
- reason
- nursery
- internet
:::

## Access the Data

```{r}
# math dataset
math <- read_delim("data/student-mat.csv", delim = ';')
# language dataset
lang <- read_delim("data/student-por.csv", delim = ';')
```

```{r}
head(math)
```

```{r}
head(lang)
```

## Clean and Combine the Data

Some ideas: 

- covert school, sex, address, famsize, Pstatus, Medu, Fedu, Mjob, Fjob, reason,
and guardian to factors
- could convert anything labeled "yes"/"no" to True or False
- eventually, drop columns that we don't need (but we wont know which ones until
we are done writing the lab, so will have to do this after lab created)

Could also combine the datasets (but this could be done in a specific question
that requires it)

## Basic Questions

... TODO ...

## Advanced Questions

Write a function to ... Apply it to both of the datasets.

```{r}
# TODO: add label
```

```{r}
#| code-fold: true
#| code-summary: "Click to see the solution"
```

## Section 4: Wrap-Up

... TODO ...
